---
title: "Consiousness - MRI in ICH Project"
output:
  html_document: default
  pdf_document: default
  word_document: default
---
# Morphological data analysis (volumetrie & location)
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r preproc, include = FALSE}
rm(list = ls())
library(readxl)
library(ggplot2)
library(dplyr)
library(reshape2)
library(doBy)


#setwd("E:/Experiments/BrainFlowMap Stuff/Ben files/MRI_ICH")

setwd("E:/Experiments/ICH_MRI/Ben files/MRI_ICH")

Loc <- read_excel("ICH_MRI_location-asr.xlsx", sheet=1)
Vol <- read_excel("Jan_Claassen_ICH_Study_Calc.xlsx", sheet=1)
Outcome <- read_excel("MRI Imaging list radiology comprehensive-br.xlsx", sheet=1)
Death <-read_excel("DeathDates.xlsx", sheet=1)

names(Loc)[names(Loc)=="Subfalcine herniation"] <- "sf_hern"
names(Loc)[names(Loc)=="MRI date"] <- "MRI_date_Loc"

names(Vol)[names(Vol)=="Hemorrhage Volume"] <- "Hg_vol"
names(Vol)[names(Vol)=="Edema Volume"] <- "Ed_vol"
names(Vol)[names(Vol)=="Total Brain Volume"] <- "Brain_vol"

names(Outcome)[names(Outcome)=="Command score before discharge"] <- "follow"
names(Outcome)[names(Outcome)=="Command Score @ time of MRI eval"] <- "MRI_Cs"

names(Death)[names(Death)=="PATIENT_ID"] <- "MRN"

# correction data
# remove duplicate death dates (one day gap, keep first)
Death[!is.na(Death$MRN) & Death$MRN==5508540,][2,]<-NA
Death[!is.na(Death$MRN) & Death$MRN==1917504,][2,]<-NA
```

## 1) Check volume data
**Note:** This need to be fixed w/ Key new data and Alex before recompute logistic regressions  
(We still have a lot of NAs because of duplicates)

#### Brain Volume:
```{r volume, echo=FALSE, warning=FALSE, fig.cap="Brain volume"}
# Control of  brain volumes  # to be fixed
Vol$Brain_vol<-as.numeric(Vol$Brain_vol)
Vol$Hg_vol<-as.numeric(Vol$Hg_vol)
Vol$Ed_vol<-as.numeric(Vol$Ed_vol)
Vol$MRN<-as.numeric(Vol$MRN)

boxplot(Vol$Brain_vol)
Vol[!is.na(Vol$Brain_vol) & Vol$Brain_vol == 0,]$Brain_vol<-NA
summary(as.numeric(Vol$Brain_vol))
```
#### Hemorage Volume
```{r, echo=FALSE, warning=FALSE, fig.cap="Hemorage volume"}
# Control of  brain volumes  # to be fixed
boxplot(Vol$Hg_vol)
Vol[!is.na(Vol$Hg_vol) & Vol$Hg_vol == 0,]$Hg_vol<-NA
summary(as.numeric(Vol$Hg_vol))
```
#### Edema Volume
```{r, echo=FALSE, warning=FALSE, fig.cap="Edema volume"}
# Control of  brain volumes  # to be fixed
boxplot(Vol$Ed_vol)
Vol[!is.na(Vol$Ed_vol) & Vol$Ed_vol == 0,]$Ed_vol<-NA
summary(as.numeric(Vol$Ed_vol))
```

```{r, echo=FALSE}
# Merging  ATTENTION here I take all Loc MRN for the anat plot (pb is that we still have duplicates MRN in Vol)

#data.raw <-merge(Loc, Vol, by.x= c("MRN"), all.x=TRUE, all.y=FALSE)
#data.raw <-merge( data.raw, Outcome, by=c("MRN"),all.x=TRUE)
#data.raw <-merge( data.raw, Death, by=c("MRN"),all.x=TRUE)

# data.raw <-merge(Loc, Outcome, by.x= c("MRN"), all.x=TRUE, all.y=FALSE)   # 
# data.raw <-merge(data.raw, Death, by.x= c("MRN"), all.x=TRUE, all.y=FALSE)   # 

data.raw <-merge(Loc, Vol, by.x= c("MRN"), all.x=TRUE, all.y=FALSE)
data.raw <-merge( data.raw, Outcome, by=c("MRN"),all.x=TRUE)
data.raw <-merge( data.raw, Death, by=c("MRN"),all.x=TRUE)

# recode errors or ambiguous data
# table (data.raw$follow); table (data.raw$MRI_Cs)
data.raw[!is.na(data.raw$follow) & data.raw$follow=='3 or 4',]$follow=3
data.raw[!is.na(data.raw$MRI_Cs) & data.raw$MRI_Cs == "4 or 5",]$MRI_Cs<- 4
```

## 2) Check consciousness data
We have 158 patients with both anat location & consciousness mesurements (@MRI & discharge)

Definition                         Score
-----------                        -------
No eyes openning                   0  
Opens eyes to stimulation          1  
Opens eyes spontaneously           2  
Tracks/attends                     3  
Follows simple commands            4  
Follows complex commands/oriented  5


Table: Consciousness ICU scale reminder.



```{r, include = FALSE}
# shape dates 
data.raw$Discharge_date<-as.Date(data.raw$Discharge_date)
data.raw$DEATH_DATE<-as.Date(data.raw$DEATH_DATE)
data.raw$MRI_date_Loc<-as.numeric(data.raw$MRI_date_Loc)
data.raw$MRI_date_Loc<-as.Date(data.raw$MRI_date_Loc,origin = "1899-12-30")

# Missed patients following to get on Eclipse (DONE)
data.raw[is.na(data.raw$follow),c("MRN","MRI_date_Loc")]
```

Patient who died in ICU:
```{r died patients, echo=FALSE}
dim(data.raw[!is.na(data.raw$DEATH_DATE) & (data.raw$Discharge_date>=data.raw$DEATH_DATE)==TRUE,])[1]

# Correction for patient who died in ICU: "following at discharge" = 0  (n=1) 
data.raw[!is.na(data.raw$DEATH_DATE) & (data.raw$Discharge_date>=data.raw$DEATH_DATE)==TRUE,]$follow<-0
```

```{r Consciousness data, echo=FALSE}
# dataframe for anat plot
loc_analyse.raw<-subset(data.raw,select=c(MRN,4,7:87,follow,MRI_Cs,Discharge_date,DEATH_DATE,MRI_date_Loc,mri_date,Hg_vol))

loc_analyse.raw$follow<-as.integer(loc_analyse.raw$follow)
loc_analyse.raw$MRI_Cs<-as.integer(loc_analyse.raw$MRI_Cs)

barplot(table(loc_analyse.raw$MRI_Cs))
title(main = list("1: Consiousness at MRI", font = 4))
table(loc_analyse.raw$MRI_Cs)
sum(table(loc_analyse.raw$MRI_Cs))

barplot(table(loc_analyse.raw$follow))
title(main = list("2: Consiousness at discharge", font = 4))
table(loc_analyse.raw$follow)
sum(table(loc_analyse.raw$follow))
```



```{r ipsi/contro, echo=FALSE}
# Add ispi / controlateral variables

#### Do the sub_tent and sus_tent checking ####
# sub_tent_names <- c("AntPons_ICH_R",
#                     "AntPons_ICH_L",
#                     "Teg_ICH_R",
#                     "Teg_ICH_L",
#                     "Cereb_ICH_R",
#                     "Cereb_ICH_L",
#                     "Vermis_ICH",
#                     "MB_ICH_C",
#                     "MB_peduncle_ICH_R",
#                     "MB_peduncle_ICH_L")
# 
# R_names <- c("TH_ant_ICH_R",
#              "TH_lat_ICH_R",
#              "TH_med_ICH_R",
#              "TH_post_ICH_R",
#              "GP_ICH_R",
#              "PUT_ICH_R",
#              "Caudate_ICH_R",
#              "IC_ant_ICH_R",
#              "IC_post_ICH_R",
#              "FCx_ICH_R",
#              "PCx_ICH_R",
#              "TCx_ICH_R",
#              "OCx_ICH_R")
# 
# L_names <- c("TH_ant_ICH_L",
#              "TH_lat_ICH_L",
#              "TH_med_ICH_L",
#              "TH_post_ICH_L",
#              "GP_ICH_L",
#              "PUT_ICH_L",
#              "Caudate_ICH_L",
#              "IC_ant_ICH_L",
#              "IC_post_ICH_L",
#              "FCx_ICH_L",
#              "PCx_ICH_L",
#              "TCx_ICH_L",
#              "OCx_ICH_L")




### Try putting all R and L from sub_tent_names into the R_names and L_names ###

sub_tent_names <- c("Vermis_ICH",
                    "MB_ICH_C")

R_names <- c("MB_peduncle_ICH_R",
             "Cereb_ICH_R",
             "Teg_ICH_R",
             "AntPons_ICH_R",
             "TH_ant_ICH_R",
             "TH_lat_ICH_R",
             "TH_med_ICH_R",
             "TH_post_ICH_R",
             "GP_ICH_R",
             "PUT_ICH_R",
             "Caudate_ICH_R",
             "IC_ant_ICH_R",
             "IC_post_ICH_R",
             "FCx_ICH_R",
             "PCx_ICH_R",
             "TCx_ICH_R",
             "OCx_ICH_R")

L_names <- c("MB_peduncle_ICH_L",
             "Cereb_ICH_L",
             "Teg_ICH_L",
             "AntPons_ICH_L",
             "TH_ant_ICH_L",
             "TH_lat_ICH_L",
             "TH_med_ICH_L",
             "TH_post_ICH_L",
             "GP_ICH_L",
             "PUT_ICH_L",
             "Caudate_ICH_L",
             "IC_ant_ICH_L",
             "IC_post_ICH_L",
             "FCx_ICH_L",
             "PCx_ICH_L",
             "TCx_ICH_L",
             "OCx_ICH_L")



# melt seperate tables containing just the MRN and the respective names, and then merge them together
# to have 3 seperate columsns for each area to check.

loc_analyse.mSub <- melt(loc_analyse.raw[,c("MRN", sub_tent_names)], id="MRN")
loc_analyse.mR <- melt(loc_analyse.raw[,c("MRN", R_names)], id="MRN")
loc_analyse.mL <- melt(loc_analyse.raw[,c("MRN", L_names)], id="MRN")

# rename the variable and value column names
colnames(loc_analyse.mSub)<-  c("MRN", "sub_tent_name", "sub_tent_value")
colnames(loc_analyse.mR)  <-  c("MRN", "R_name", "R_value")
colnames(loc_analyse.mL)  <-  c("MRN", "L_name", "L_value")

# now merge.
# This creates "duplicate" rows based on all the various combos you can (and have to) make
# between all the Sub, R, and L names
# but that shouldn't matter, because we're just checking if a 1 exists in any of these columns
# per patient, and that shouldn't change even if rows are "duplicated"
loc_analyse.m <- merge(loc_analyse.mSub, merge(loc_analyse.mL, loc_analyse.mR))

# create two new columns, one for the sub_tent stuff and another for the L, R, Both, None stuff
# then check each column for each patient and assign to these new columns accordingly

loc_analyse.raw2<-loc_analyse.raw
loc_analyse.raw2$sub_tent <- NA
loc_analyse.raw2$sus_tent <- NA

# find unique patient IDs to loop over
ID.u <- unique(loc_analyse.raw2$MRN)

# check via looping over each MRN
# I think I need to do this without dplyr because
# that way, when filtering the table, I can keep the same indices of those filtered rows
# when assigning the value of 'sus_tentorial' back into the main table.

for (i in 1:length(ID.u)) {
  tmp <- loc_analyse.m[loc_analyse.m$MRN == ID.u[i], ]
  
  # check for sub_tent first
  if (any(tmp$sub_tent_value >= 1)) {
    loc_analyse.raw2[loc_analyse.raw2$MRN == ID.u[i], ]$sub_tent <- 1
  } else {
    loc_analyse.raw2[loc_analyse.raw2$MRN == ID.u[i], ]$sub_tent <- 0
  }
  
  
  # now check for the L, R, Both, None
  if (any(tmp$L_value >= 1) & !any(tmp$R_value >= 1)) {  # if patient has lesion on left side and not right side
    loc_analyse.raw2[loc_analyse.raw2$MRN == ID.u[i],]$sus_tent <- "L"
    
  } else if (!any(tmp$L_value >= 1) & any(tmp$R_value >= 1)) {  # if patient doesn't have lesion of left side but has on right side
    loc_analyse.raw2[loc_analyse.raw2$MRN == ID.u[i],]$sus_tent <- "R"
    
  } else if (any(tmp$L_value >= 1) & any(tmp$R_value >= 1)) {  # if patient has lesion on both sides
    loc_analyse.raw2[loc_analyse.raw2$MRN == ID.u[i],]$sus_tent <- "Both"
    
  } else if (!any(tmp$L_value >= 1) & !any(tmp$R_value >= 1)) {  # if patient doesn't have lesion on either side
    loc_analyse.raw2[loc_analyse.raw2$MRN == ID.u[i],]$sus_tent <- "None"
    
  }
}


### So now loc_analyse.m contains info, for each patient, on whether or not
### they had a lesion in the sub_tent area or if they had a lesion in the sus_tent area
### on the left, right, both, or none of these sides.

## BEN: Kevin I changed minimally your script to get one line / patient (filling loc_analyse.raw2) 
```


```{r recode following scale, echo=FALSE, warning=FALSE}

loc_analyse.raw<- loc_analyse.raw2;
# now loc_analyse.raw has the 2 more row sub_tent & sus_tent


# recode following scale
# in binary in xxx2
Cs_cutoff=3
loc_analyse.raw$follow2<-loc_analyse.raw$follow
loc_analyse.raw[loc_analyse.raw$follow2 < Cs_cutoff,]$follow2<- 0
loc_analyse.raw[loc_analyse.raw$follow2 >= Cs_cutoff,]$follow2<- 1
loc_analyse.raw$MRI_Cs2<- loc_analyse.raw$MRI_Cs
loc_analyse.raw[!is.na(loc_analyse.raw$MRI_Cs2) & loc_analyse.raw$MRI_Cs2 <Cs_cutoff,]$MRI_Cs2<- 0
loc_analyse.raw[!is.na(loc_analyse.raw$MRI_Cs2) & loc_analyse.raw$MRI_Cs2 >=Cs_cutoff,]$MRI_Cs2<- 1

# in 3 groups in xxx3
loc_analyse.raw$follow3<-loc_analyse.raw$follow
loc_analyse.raw[loc_analyse.raw$follow3 < 1,]$follow3<- 0
loc_analyse.raw[(loc_analyse.raw$follow3 >= 1) & (loc_analyse.raw$follow3 <4)  ,]$follow3<- 1
loc_analyse.raw[loc_analyse.raw$follow3 >= 4 ,]$follow3<- 2
loc_analyse.raw$MRI_Cs3<- loc_analyse.raw$MRI_Cs
loc_analyse.raw[!is.na(loc_analyse.raw$MRI_Cs3) & loc_analyse.raw$MRI_Cs3 <1,]$MRI_Cs3<- 0
loc_analyse.raw[!is.na(loc_analyse.raw$MRI_Cs3) & (loc_analyse.raw$MRI_Cs3 >= 1) & (loc_analyse.raw$MRI_Cs3 <4) ,]$MRI_Cs3<- 1
loc_analyse.raw[!is.na(loc_analyse.raw$MRI_Cs3) & loc_analyse.raw$MRI_Cs3 >=4,]$MRI_Cs3<- 2

# plot with original 6 cat scale done in "plot consciousness data"
```

#### Split in 2 categories: 0 = No attending/traking / 1 = Attending/traking or better
```{r, echo=FALSE}
# plot with 2 cat
barplot(table(loc_analyse.raw$MRI_Cs2))
title(main = list("3: Consiousness at MRI", font = 4))
barplot(table(loc_analyse.raw$follow2))
title(main = list("4: Consiousness at discharge", font = 4))
table(loc_analyse.raw$MRI_Cs2); table(loc_analyse.raw$follow2)
```

#### Split in 3 categories: 0 = Coma / 1 = "Awakening" (1,2,3) / 2 = Following commands (4,5)
```{r, echo=FALSE,warning = FALSE}
# plot with 3 cat
barplot(table(loc_analyse.raw$MRI_Cs3))
title(main = list("5: Consiousness at MRI", font = 4))
barplot(table(loc_analyse.raw$follow3))
title(main = list("6: Consiousness at discharge", font = 4))
table(loc_analyse.raw$MRI_Cs3); table(loc_analyse.raw$follow3)
sum(table(loc_analyse.raw$MRI_Cs3))


# recodin cortex values >1 as 1
loc_analyse.raw[,2:76]<- ifelse(loc_analyse.raw[,2:76]>=1,1,0)
loc_analyse.raw[,2:76]<- lapply(loc_analyse.raw[,2:76], as.factor)

# rm NAs
#loc_analyse.raw <- loc_analyse.raw[complete.cases(loc_analyse.raw[,2:75]),] No need here

```

```{r, echo=FALSE,warning = FALSE}

# adding new variables

# Thalamus Thal_R Thal_L
loc_analyse.raw$TH_ICH_L <- NA
loc_analyse.raw$TH_ICH_R <- NA

loc_analyse.raw$TH_ICH_L <- ifelse(loc_analyse.raw[,"TH_ant_ICH_L"]==1 |loc_analyse.raw[,"TH_med_ICH_L"]==1|loc_analyse.raw[,"TH_lat_ICH_L"]==1 |loc_analyse.raw[,"TH_post_ICH_L"]==1 ,1,0)

loc_analyse.raw$TH_ICH_R <- ifelse(loc_analyse.raw[,"TH_ant_ICH_R"]==1 |loc_analyse.raw[,"TH_med_ICH_R"]==1|loc_analyse.raw[,"TH_lat_ICH_R"]==1 |loc_analyse.raw[,"TH_post_ICH_R"]==1 ,1,0)

loc_analyse.raw$TH_edema_L <- NA
loc_analyse.raw$TH_edema_R <- NA

loc_analyse.raw$TH_edema_L <- ifelse(loc_analyse.raw[,"TH_ant_edema_L"]==1 |loc_analyse.raw[,"TH_med_edema_L"]==1|loc_analyse.raw[,"TH_lat_edema_L"]==1 |loc_analyse.raw[,"TH_post_edema_L"]==1 ,1,0)

loc_analyse.raw$TH_edema_R <- ifelse(loc_analyse.raw[,"TH_ant_edema_R"]==1 |loc_analyse.raw[,"TH_med_edema_R"]==1|loc_analyse.raw[,"TH_lat_edema_R"]==1 |loc_analyse.raw[,"TH_post_edema_R"]==1 ,1,0)


# Mesocicuite MesoC_R MesoC_L MesoC
loc_analyse.raw$MesoC_ICH_R <- NA
loc_analyse.raw$MesoC_ICH_L <- NA
loc_analyse.raw$MesoC_ICH_Bi <- NA
loc_analyse.raw$MesoC_ICH_Uni <- NA

loc_analyse.raw$MesoC_ICH_R <- ifelse(loc_analyse.raw[,"FCx_ICH_R"]==1|
                                  loc_analyse.raw[,"TH_ICH_R"]==1|
                                  loc_analyse.raw[,"GP_ICH_R"]==1|
                                  loc_analyse.raw[,"Caudate_ICH_R"]==1|
                                  loc_analyse.raw[,"PUT_ICH_R"]==1|
                                  loc_analyse.raw[,"MB_peduncle_ICH_R"]==1|
                                  loc_analyse.raw[,"MB_ICH_C"]==1|
                                  loc_analyse.raw[,"Teg_ICH_R"]==1 ,1,0)

loc_analyse.raw$MesoC_ICH_L <- ifelse(loc_analyse.raw[,"FCx_ICH_L"]==1|
                                  loc_analyse.raw[,"TH_ICH_L"]==1|
                                  loc_analyse.raw[,"GP_ICH_L"]==1|
                                  loc_analyse.raw[,"Caudate_ICH_L"]==1|
                                  loc_analyse.raw[,"PUT_ICH_L"]==1|
                                  loc_analyse.raw[,"MB_peduncle_ICH_L"]==1|
                                  loc_analyse.raw[,"MB_ICH_C"]==1|
                                  loc_analyse.raw[,"Teg_ICH_L"]==1 ,1,0)

loc_analyse.raw$MesoC_ICH_Bi<- ifelse(loc_analyse.raw[,"MesoC_ICH_R"]==1 &
                                  loc_analyse.raw[,"MesoC_ICH_L"]==1 ,1,0)

loc_analyse.raw$MesoC_ICH_Uni<- ifelse(loc_analyse.raw[,"MesoC_ICH_R"]==1 |
                                  loc_analyse.raw[,"MesoC_ICH_L"]==1 ,1,0)

# Brainstem
loc_analyse.raw$BS_ICH <- NA 
loc_analyse.raw$BS_ICH <- ifelse(loc_analyse.raw[,"MB_peduncle_ICH_R"]==1|
                              loc_analyse.raw[,"MB_peduncle_ICH_L"]==1|
                              loc_analyse.raw[,"MB_ICH_C"]==1|
                              loc_analyse.raw[,"Teg_ICH_R"]==1|
                              loc_analyse.raw[,"Teg_ICH_L"]==1 ,1,0)

```

**Note:** MesoC = TH + GP + Caudate + PUT + Teg + MB_pedoncle + MB_C + FCx
I took only ICH yet


```{r, echo=FALSE,warning = FALSE}
#  melting and using dplyr  (thanks Kevin)
dat.m <- melt(loc_analyse.raw, id=c("MRN", "Discharge_date", "DEATH_DATE", "MRI_date_Loc", "mri_date", "follow2","MRI_Cs2", "follow3","MRI_Cs3"))

# ERROR here why?
### This happens because some of the columns being melted are factors with different levels from one column to another.
### When it tries to melt these columns, it sees that the factor levels aren't the same between them, so it apparently
### just drops the levels and converts everything to character.
### So, basically, I don't think we really need to care about this warning.
# BEN: OK

dat.m$value<-as.numeric(dat.m$value)
```

## 3) Plot anatomical data 
Percentage (not mean as ploted) of patient with a lesion (ICH or eadema) in each ROIs according to level of consciousness (at time of MRI or discharge))

### Consiousness split in 2 categories
#### Right / Left analyse

```{r, echo=FALSE}
# for follow at MRI
avg2<- dat.m %>% 
  group_by(MRI_Cs2, variable) %>%
  filter( !(variable  %in%  c("Cerebellar tonsillar herniation","MLS [mm]", "Old stroke", "Old ICH",  "Uncal herniation (to which side) ","Transtentorial herniation ","Uncal herniation (to which side)","Transtentorial herniation","sf_hern", "Hypo_edema_C","Hypo_ICH_C","TH_ant_ICH_L","TH_med_ICH_L","TH_lat_ICH_L","TH_post_ICH_L",
"TH_ant_ICH_R","TH_med_ICH_R","TH_lat_ICH_R","TH_post_ICH_R",
"TH_ant_edema_L","TH_med_edema_L","TH_lat_edema_L","TH_post_edema_L", 
"TH_ant_edema_R","TH_med_edema_R","TH_lat_edema_R","TH_post_edema_R") )) %>%
  summarize(mean=mean(value,na.rm=TRUE))

avg2$MRI_Cs2<-as.integer(avg2$MRI_Cs2);avg2$MRI_Cs2<-as.factor(avg2$MRI_Cs2)


ggplot(data=na.omit(rbind(avg2[grep(pattern = "ICH", avg2$variable),],avg2[avg2$variable=="IVH",])), aes(x=variable, y=mean, fill=MRI_Cs2) )+
  geom_bar(stat="identity", position=position_dodge()) +
  coord_flip() +
  labs(title = "7: ICH ~ Follow @ mri")

ggplot(data=na.omit(avg2[grep(pattern = "edema", avg2$variable),  ]), aes(x=variable, y=mean, fill=MRI_Cs2) )+
  geom_bar(stat="identity", position=position_dodge()) +
  coord_flip()+
  labs(title = "8: Edema ~ Follow @ mri")


# for follow at discharge
avg1<- dat.m %>% 
  group_by(follow2, variable) %>%
  filter( !(variable  %in%  c("Cerebellar tonsillar herniation","MLS [mm]", "Old stroke", "Old ICH",  "Uncal herniation (to which side) ","Transtentorial herniation ",
                              "Uncal herniation (to which side)","Transtentorial herniation","sf_hern", "Hypo_edema_C","Hypo_ICH_C","TH_ant_ICH_L","TH_med_ICH_L","TH_lat_ICH_L","TH_post_ICH_L",
"TH_ant_ICH_R","TH_med_ICH_R","TH_lat_ICH_R","TH_post_ICH_R",
"TH_ant_edema_L","TH_med_edema_L","TH_lat_edema_L","TH_post_edema_L", 
"TH_ant_edema_R","TH_med_edema_R","TH_lat_edema_R","TH_post_edema_R") )) %>%
  #summarize(mean=mean(value,na.rm=TRUE), sd=sd(value,na.rm=TRUE)) no need sd (%)
  summarize(mean=mean(value,na.rm=TRUE))
  

avg1$follow2<-as.integer(avg1$follow2);avg1$follow2<-as.factor(avg1$follow2)

ggplot(data=rbind(avg1[grep(pattern = "ICH", avg1$variable),],avg1[avg1$variable=="IVH",]), aes(x=variable, y=mean, fill=follow2) )+
  geom_bar(stat="identity", position=position_dodge()) +
  coord_flip() +
  labs(title = "9: ICH ~ Follow @ discharge")

ggplot(data=avg1[grep(pattern = "edema", avg1$variable),  ], aes(x=variable, y=mean, fill=follow2) )+
  geom_bar(stat="identity", position=position_dodge()) +
  coord_flip()+
  labs(title = "10: Edema ~ Follow @ discharge")


```


#### Ipsi / Controlateral analyse

In the following table, left 0/1 code corresponds to infra-tentorial ICH and R/L.. to supra-tentorial ICH  
**Note:** We have one patients without infra or supra tent ICH or IVH (check w/ Alex)

```{r}
table(loc_analyse.raw2$sub_tent,loc_analyse.raw2$sus_tent)
sum(table(loc_analyse.raw2$sub_tent,loc_analyse.raw2$sus_tent))

# patients without any hemorrage ??
loc_analyse.raw2[loc_analyse.raw2$IVH==0 & loc_analyse.raw2$sub_tent==0 & loc_analyse.raw2$sus_tent=="None" ,]$MRN
```

```{r}
#n=
62+7+52+5
```


```{r, echo=FALSE, warning=FALSE}
## analyse ipsi / contro
loc_analyse.raw3<-loc_analyse.raw[loc_analyse.raw$sus_tent=='L',]

# R->contro L->ipsi

colnames(loc_analyse.raw3) <- c(
  "MRN", "IVH", "AntPons_edema_ipsi", "AntPons_edema_contro",    
  "AntPons_ICH_ipsi",   "AntPons_ICH_contro",   "Teg_edema_ipsi",    
   "Teg_edema_contro",     "Teg_ICH_ipsi",       "Teg_ICH_contro",      
   "MB_edema_C",      "MB_peduncle_edema_ipsi", "MB_peduncle_edema_contro",    
   "MB_ICH_C",        "MB_peduncle_ICH_ipsi",   "MB_peduncle_ICH_contro",  
   "TH_ant_edema_ipsi",  "TH_ant_edema_contro",  "TH_lat_edema_ipsi", 
   "TH_lat_edema_contro",  "TH_med_edema_ipsi",  "TH_med_edema_contro", 
   "TH_post_edema_ipsi", "TH_post_edema_contro", "TH_ant_ICH_ipsi",   
   "TH_ant_ICH_contro",    "TH_lat_ICH_ipsi",    "TH_lat_ICH_contro",   
   "TH_med_ICH_ipsi",    "TH_med_ICH_contro",    "TH_post_ICH_ipsi",  
   "TH_post_ICH_contro",   "GP_edema_ipsi",      "GP_edema_contro",     
   "PUT_edema_ipsi",     "PUT_edema_contro",     "Caudate_edema_ipsi",    
   "Caudate_edema_contro", "IC_ant_edema_ipsi",  "IC_ant_edema_contro", 
   "IC_post_edema_ipsi", "IC_post_edema_contro", "GP_ICH_ipsi",       
   "GP_ICH_contro",        "PUT_ICH_ipsi",       "PUT_ICH_contro",      
   "Caudate_ICH_ipsi",   "Caudate_ICH_contro",   "IC_ant_ICH_ipsi",   
   "IC_ant_ICH_contro",    "IC_post_ICH_ipsi",   "IC_post_ICH_contro",  
   "FCx_edema_ipsi",     "FCx_edema_contro",     "Hypo_edema_C",   
   "FCx_ICH_ipsi",       "FCx_ICH_contro",       "Hypo_ICH_C",     
   "PCx_edema_ipsi",     "PCx_edema_contro",     "PCx_ICH_ipsi",      
   "PCx_ICH_contro",       "TCx_edema_ipsi",     "TCx_edema_contro",    
   "TCx_ICH_ipsi",       "TCx_ICH_contro",       "OCx_edema_ipsi",    
   "OCx_edema_contro",     "OCx_ICH_ipsi",       "OCx_ICH_contro",      
   "Cereb_edema_ipsi",   "Cereb_edema_contro",   "Vermis_edema",   
   "Cereb_ICH_ipsi",     "Cereb_ICH_contro",     "Vermis_ICH",     
   "sf_hern", "Uncal herniation (to which side)","Transtentorial herniation",      
   "Cerebellar tonsillar herniation", "MLS [mm]",        "Old stroke",     
   "Old ICH", "follow",  "MRI_Cs", 
   "Discharge_date",  "DEATH_DATE",      "MRI_date_Loc",   
   "mri_date",        "sub_tent",        "sus_tent",       
   "follow2", "MRI_Cs2", "follow3",        
   "MRI_Cs3", "TH_ICH_ipsi", "TH_ICH_contro" , "TH_edema_ipsi", "TH_edema_contro",
  "MesoC_ICH_contro", "MesoC_ICH_ipsi", "MesoC_ICH_Bi", "MesoC_ICH_Uni", "BS_ICH" )

# R->contro L->ipsi

temp<-loc_analyse.raw[loc_analyse.raw$sus_tent=='R',]

# L->contro R->ipsi

colnames(temp) <- c(
  "MRN", "IVH","AntPons_edema_contro", "AntPons_edema_ipsi",    
  "AntPons_ICH_contro",   "AntPons_ICH_ipsi",   "Teg_edema_contro",    
  "Teg_edema_ipsi",     "Teg_ICH_contro",       "Teg_ICH_ipsi",      
  "MB_edema_C",      "MB_peduncle_edema_contro", "MB_peduncle_edema_ipsi",    
  "MB_ICH_C",        "MB_peduncle_ICH_contro",   "MB_peduncle_ICH_ipsi",  
  "TH_ant_edema_contro",  "TH_ant_edema_ipsi",  "TH_lat_edema_contro", 
  "TH_lat_edema_ipsi",  "TH_med_edema_contro",  "TH_med_edema_ipsi", 
  "TH_post_edema_contro", "TH_post_edema_ipsi", "TH_ant_ICH_contro",   
  "TH_ant_ICH_ipsi",    "TH_lat_ICH_contro",    "TH_lat_ICH_ipsi",   
  "TH_med_ICH_contro",    "TH_med_ICH_ipsi",    "TH_post_ICH_contro",  
  "TH_post_ICH_ipsi",   "GP_edema_contro",      "GP_edema_ipsi",     
  "PUT_edema_contro",     "PUT_edema_ipsi",     "Caudate_edema_contro",    
  "Caudate_edema_ipsi", "IC_ant_edema_contro",  "IC_ant_edema_ipsi", 
  "IC_post_edema_contro", "IC_post_edema_ipsi", "GP_ICH_contro",       
  "GP_ICH_ipsi",        "PUT_ICH_contro",       "PUT_ICH_ipsi",      
  "Caudate_ICH_contro",   "Caudate_ICH_ipsi",   "IC_ant_ICH_contro",   
  "IC_ant_ICH_ipsi",    "IC_post_ICH_contro",   "IC_post_ICH_ipsi",  
  "FCx_edema_contro",     "FCx_edema_ipsi",     "Hypo_edema_C",   
  "FCx_ICH_contro",       "FCx_ICH_ipsi",       "Hypo_ICH_C",     
  "PCx_edema_contro",     "PCx_edema_ipsi",     "PCx_ICH_contro",      
  "PCx_ICH_ipsi",       "TCx_edema_contro",     "TCx_edema_ipsi",    
  "TCx_ICH_contro",       "TCx_ICH_ipsi",       "OCx_edema_contro",    
  "OCx_edema_ipsi",     "OCx_ICH_contro",       "OCx_ICH_ipsi",      
  "Cereb_edema_contro",   "Cereb_edema_ipsi",   "Vermis_edema",   
  "Cereb_ICH_contro",     "Cereb_ICH_ipsi",     "Vermis_ICH",     
  "sf_hern", "Uncal herniation (to which side)","Transtentorial herniation",      
  "Cerebellar tonsillar herniation", "MLS [mm]",        "Old stroke",     
  "Old ICH", "follow",  "MRI_Cs", 
  "Discharge_date",  "DEATH_DATE",      "MRI_date_Loc",   
  "mri_date",        "sub_tent",        "sus_tent",       
  "follow2", "MRI_Cs2", "follow3",        
  "MRI_Cs3", "TH_ICH_contro", "TH_ICH_ipsi" , "TH_edema_contro", "TH_edema_ipsi",
  "MesoC_ICH_ipsi", "MesoC_ICH_contro", "MesoC_ICH_Bi", "MesoC_ICH_Uni", "BS_ICH" )

# L->contro R->ipsi

loc_analyse.raw3<-rbind(temp,loc_analyse.raw3)


dat2.m <- melt(loc_analyse.raw3, id=c("MRN", "Discharge_date", "DEATH_DATE", "MRI_date_Loc", "mri_date", "follow2","MRI_Cs2", "follow3","MRI_Cs3"))
dat2.m$value<-as.numeric(dat2.m$value)

# for follow at MRI
avg2<- dat2.m %>% 
  group_by(MRI_Cs2, variable) %>%
  filter( !(variable  %in%  c("Cerebellar tonsillar herniation","MLS [mm]", "Old stroke", "Old ICH",  "Uncal herniation (to which side) ","Transtentorial herniation ",
                              "Uncal herniation (to which side)","Transtentorial herniation","sf_hern", "Hypo_edema_C","Hypo_ICH_C","TH_ant_edema_contro",  "TH_ant_edema_ipsi",  "TH_lat_edema_contro", 
  "TH_lat_edema_ipsi",  "TH_med_edema_contro",  "TH_med_edema_ipsi", 
  "TH_post_edema_contro", "TH_post_edema_ipsi", "TH_ant_ICH_contro",   
  "TH_ant_ICH_ipsi",    "TH_lat_ICH_contro",    "TH_lat_ICH_ipsi",   
  "TH_med_ICH_contro",    "TH_med_ICH_ipsi",    "TH_post_ICH_contro",  
  "TH_post_ICH_ipsi") )) %>%
  summarize(mean=mean(value,na.rm=TRUE))

avg2$MRI_Cs2<-as.integer(avg2$MRI_Cs2);avg2$MRI_Cs2<-as.factor(avg2$MRI_Cs2)

ggplot(data=na.omit(rbind(avg2[grep(pattern = "ICH", avg2$variable),  ],avg2[avg2$variable=="IVH",])), aes(x=variable, y=mean, fill=MRI_Cs2) )+
  geom_bar(stat="identity", position=position_dodge()) +
  coord_flip() +
  labs(title = "11: ICH ~ Follow @ mri")

ggplot(data=na.omit(avg2[grep(pattern = "edema", avg2$variable),  ]), aes(x=variable, y=mean, fill=MRI_Cs2) )+
  geom_bar(stat="identity", position=position_dodge()) +
  coord_flip()+
  labs(title = "12: Edema ~ Follow @ mri")



# for follow at discharge
avg1<- dat2.m %>% 
  group_by(follow2, variable) %>%
  filter( !(variable  %in%  c("Cerebellar tonsillar herniation","MLS [mm]", "Old stroke", "Old ICH",  "Uncal herniation (to which side) ","Transtentorial herniation ",
                              "Uncal herniation (to which side)","Transtentorial herniation","sf_hern", "Hypo_edema_C","Hypo_ICH_C","TH_ant_edema_contro",  "TH_ant_edema_ipsi",  "TH_lat_edema_contro", 
  "TH_lat_edema_ipsi",  "TH_med_edema_contro",  "TH_med_edema_ipsi", 
  "TH_post_edema_contro", "TH_post_edema_ipsi", "TH_ant_ICH_contro",   
  "TH_ant_ICH_ipsi",    "TH_lat_ICH_contro",    "TH_lat_ICH_ipsi",   
  "TH_med_ICH_contro",    "TH_med_ICH_ipsi",    "TH_post_ICH_contro",  
  "TH_post_ICH_ipsi") )) %>%
  #summarize(mean=mean(value,na.rm=TRUE), sd=sd(value,na.rm=TRUE)) no need sd (%)
  summarize(mean=mean(value,na.rm=TRUE))


avg1$follow2<-as.integer(avg1$follow2);avg1$follow2<-as.factor(avg1$follow2)

ggplot(data=rbind(avg1[grep(pattern = "ICH", avg1$variable),  ],avg1[avg1$variable=="IVH",]), aes(x=variable, y=mean, fill=follow2) )+
  geom_bar(stat="identity", position=position_dodge()) +
  coord_flip()+
  labs(title = "13: ICH ~ Follow @ discharge")

ggplot(data=avg1[grep(pattern = "edema", avg1$variable),  ], aes(x=variable, y=mean, fill=follow2) )+
  geom_bar(stat="identity", position=position_dodge()) +
  coord_flip()+
  labs(title = "14: Edema ~ Follow @ discharge")

```


### Consiousness split in 3 categories
#### according to Right / Left
```{r, echo=FALSE}
# for follow at MRI
avg2<- dat.m %>% 
  group_by(MRI_Cs3, variable) %>%
  filter( !(variable  %in%  c("Cerebellar tonsillar herniation","MLS [mm]", "Old stroke", "Old ICH",  "Uncal herniation (to which side) ","Transtentorial herniation ",
                              "Uncal herniation (to which side)","Transtentorial herniation","sf_hern", "Hypo_edema_C","Hypo_ICH_C","TH_ant_edema_contro",  "TH_ant_edema_ipsi",  "TH_lat_edema_contro", 
  "TH_lat_edema_ipsi",  "TH_med_edema_contro",  "TH_med_edema_ipsi", 
  "TH_post_edema_contro", "TH_post_edema_ipsi", "TH_ant_ICH_contro",   
  "TH_ant_ICH_ipsi",    "TH_lat_ICH_contro",    "TH_lat_ICH_ipsi",   
  "TH_med_ICH_contro",    "TH_med_ICH_ipsi",    "TH_post_ICH_contro",  
  "TH_post_ICH_ipsi") )) %>%
  summarize(mean=mean(value,na.rm=TRUE))

avg2$MRI_Cs3<-as.integer(avg2$MRI_Cs3);avg2$MRI_Cs3<-as.factor(avg2$MRI_Cs3)



ggplot(data=na.omit(rbind(avg2[grep(pattern = "ICH", avg2$variable),  ],avg2[avg2$variable=="IVH",])), aes(x=variable, y=mean, fill=MRI_Cs3) )+
  geom_bar(stat="identity", position=position_dodge()) +
  coord_flip() +
  labs(title = "15: ICH ~ Follow @ mri")

ggplot(data=na.omit(avg2[grep(pattern = "edema", avg2$variable),  ]), aes(x=variable, y=mean, fill=MRI_Cs3) )+
  geom_bar(stat="identity", position=position_dodge()) +
  coord_flip()+
  labs(title = "16: Edema ~ Follow @ mri")


# for follow at discharge
avg1<- dat.m %>% 
  group_by(follow3, variable) %>%
  filter( !(variable  %in%  c("Cerebellar tonsillar herniation","MLS [mm]", "Old stroke", "Old ICH",  "Uncal herniation (to which side) ","Transtentorial herniation ",
                                "Uncal herniation (to which side)","Transtentorial herniation","sf_hern", "Hypo_edema_C","Hypo_ICH_C","TH_ant_edema_contro",  "TH_ant_edema_ipsi",  "TH_lat_edema_contro", 
  "TH_lat_edema_ipsi",  "TH_med_edema_contro",  "TH_med_edema_ipsi", 
  "TH_post_edema_contro", "TH_post_edema_ipsi", "TH_ant_ICH_contro",   
  "TH_ant_ICH_ipsi",    "TH_lat_ICH_contro",    "TH_lat_ICH_ipsi",   
  "TH_med_ICH_contro",    "TH_med_ICH_ipsi",    "TH_post_ICH_contro",  
  "TH_post_ICH_ipsi") )) %>%
  #summarize(mean=mean(value,na.rm=TRUE), sd=sd(value,na.rm=TRUE)) no need sd (%)
  summarize(mean=mean(value,na.rm=TRUE))
  

avg1$follow3<-as.integer(avg1$follow3);avg1$follow3<-as.factor(avg1$follow3)

ggplot(data=rbind(avg1[grep(pattern = "ICH", avg1$variable),  ],avg1[avg1$variable=="IVH",]), aes(x=variable, y=mean, fill=follow3) )+
  geom_bar(stat="identity", position=position_dodge()) +
  coord_flip() +
  labs(title = "17: ICH ~ Follow @ discharge")

ggplot(data=avg1[grep(pattern = "edema", avg1$variable),  ], aes(x=variable, y=mean, fill=follow3) )+
  geom_bar(stat="identity", position=position_dodge()) +
  coord_flip()+
  labs(title = "18: Edema ~ Follow @ discharge")


```


#### according to Ipsi / Controlateral
```{r, echo=FALSE, warning=FALSE}
## analyse ipsi / contro
library(cumplyr)
library(scales)
library(RColorBrewer)

table(loc_analyse.raw3$MRI_Cs3)
sum(table(loc_analyse.raw3$MRI_Cs3))

# for follow at MRI
avg2<- dat2.m %>% 
  group_by(MRI_Cs3, variable) %>%
  filter( !(variable  %in%  c("Cerebellar tonsillar herniation","MLS [mm]", "Old stroke", "Old ICH",  "Uncal herniation (to which side) ","Transtentorial herniation ",
                                "Uncal herniation (to which side)","Transtentorial herniation","sf_hern", "Hypo_edema_C","Hypo_ICH_C","TH_ant_edema_contro",  "TH_ant_edema_ipsi",  "TH_lat_edema_contro", 
  "TH_lat_edema_ipsi",  "TH_med_edema_contro",  "TH_med_edema_ipsi", 
  "TH_post_edema_contro", "TH_post_edema_ipsi", "TH_ant_ICH_contro",   
  "TH_ant_ICH_ipsi",    "TH_lat_ICH_contro",    "TH_lat_ICH_ipsi",   
  "TH_med_ICH_contro",    "TH_med_ICH_ipsi",    "TH_post_ICH_contro",  
  "TH_post_ICH_ipsi") )) %>%
  summarize(mean=mean(value,na.rm=TRUE))

avg2$MRI_Cs3<-as.integer(avg2$MRI_Cs3);avg2$MRI_Cs3<-as.factor(avg2$MRI_Cs3)

ggplot(data=na.omit(rbind(avg2[grep(pattern = "ICH", avg2$variable),],avg2[avg2$variable=="IVH",])), aes(x=variable, y=mean, fill=MRI_Cs3) )+
  geom_bar(stat="identity", position=position_dodge()) +
  coord_flip() +
  labs(title = "19: ICH ~ Follow @ mri")

ggplot( data=na.omit(rbind(avg2[grep(pattern = "ICH", avg2$variable),],avg2[avg2$variable=="IVH",]))) + 
          geom_tile(aes(x=variable, y=MRI_Cs3, fill=mean))+
          scale_fill_gradientn(colours=rev(brewer.pal(10,"Spectral")), limits=c(0,1))+
          coord_flip()+
          ggtitle("Plot for Sketch 19_2: ICH ~ Follow @ mri")

ggplot(data=na.omit(avg2[grep(pattern = "edema", avg2$variable),  ]), aes(x=variable, y=mean, fill=MRI_Cs3) )+
  geom_bar(stat="identity", position=position_dodge()) +
  coord_flip()+
  labs(title = "20: Edema ~ Follow @ mri")

# for follow at discharge
avg1<- dat2.m %>% 
  group_by(follow3, variable) %>%
  filter( !(variable  %in%  c("Cerebellar tonsillar herniation","MLS [mm]", "Old stroke", "Old ICH",  "Uncal herniation (to which side) ","Transtentorial herniation ",
                                "Uncal herniation (to which side)","Transtentorial herniation","sf_hern", "Hypo_edema_C","Hypo_ICH_C","TH_ant_edema_contro",  "TH_ant_edema_ipsi",  "TH_lat_edema_contro", 
  "TH_lat_edema_ipsi",  "TH_med_edema_contro",  "TH_med_edema_ipsi", 
  "TH_post_edema_contro", "TH_post_edema_ipsi", "TH_ant_ICH_contro",   
  "TH_ant_ICH_ipsi",    "TH_lat_ICH_contro",    "TH_lat_ICH_ipsi",   
  "TH_med_ICH_contro",    "TH_med_ICH_ipsi",    "TH_post_ICH_contro",  
  "TH_post_ICH_ipsi") )) %>%
  #summarize(mean=mean(value,na.rm=TRUE), sd=sd(value,na.rm=TRUE)) no need sd (%)
  summarize(mean=mean(value,na.rm=TRUE))
  
avg1$follow3<-as.integer(avg1$follow3);avg1$follow3<-as.factor(avg1$follow3)

ggplot(data=na.omit(rbind(avg1[grep(pattern = "ICH", avg1$variable),  ],avg1[avg1$variable=="IVH",])), aes(x=variable, y=mean, fill=follow3) )+
  geom_bar(stat="identity", position=position_dodge()) +
  coord_flip()+
  labs(title = "21: ICH ~ Follow @ discharge")

ggplot( data=na.omit(rbind(avg1[grep(pattern = "ICH", avg1$variable),  ],avg1[avg1$variable=="IVH",]))) + 
          geom_tile(aes(x=variable, y=follow3, fill=mean))+
          scale_fill_gradientn(colours=rev(brewer.pal(10,"Spectral")), limits=c(0,1))+
          coord_flip()+
          ggtitle("Plot for Sketch 21_2: ICH ~ Follow @ mri")


ggplot(data=avg1[grep(pattern = "edema", avg1$variable),  ], aes(x=variable, y=mean, fill=follow3) )+
  geom_bar(stat="identity", position=position_dodge()) +
  coord_flip()+
  labs(title = "22: Edema ~ Follow @ discharge")

```



#### subgroup of patient with bilateral sub-tentorial hemorrage (n=9)

```{r, echo=FALSE, warning=FALSE}
## analyse ipsi / contro

dat3.m <- melt(loc_analyse.raw[loc_analyse.raw$sus_tent=='Both',], id=c("MRN", "Discharge_date", "DEATH_DATE", "MRI_date_Loc", "mri_date", "follow2","MRI_Cs2", "follow3","MRI_Cs3"))
dat3.m$value<-as.numeric(dat3.m$value)

# for follow at MRI
avg2<- dat3.m %>% 
  group_by(MRI_Cs2, variable) %>%
  filter( !(variable  %in%  c("Cerebellar tonsillar herniation","MLS [mm]", "Old stroke", "Old ICH",  "Uncal herniation (to which side) ","Transtentorial herniation ",
                              "Uncal herniation (to which side)","Transtentorial herniation","sf_hern", "Hypo_edema_C","Hypo_ICH_C") )) %>%
  summarize(mean=mean(value,na.rm=TRUE))

table(loc_analyse.raw[loc_analyse.raw$sus_tent=='Both',]$MRI_Cs2)

avg2$MRI_Cs2<-as.integer(avg2$MRI_Cs2);avg2$MRI_Cs2<-as.factor(avg2$MRI_Cs2)

ggplot(data=na.omit(avg2[grep(pattern = "edema", avg2$variable),  ]), aes(x=variable, y=mean, fill=MRI_Cs2) )+
  geom_bar(stat="identity", position=position_dodge()) +
  coord_flip()+
  labs(title = "23: Edema ~ Follow @ mri")

ggplot(data=na.omit(avg2[grep(pattern = "ICH", avg2$variable),]), aes(x=variable, y=mean, fill=MRI_Cs2) )+
  geom_bar(stat="identity", position=position_dodge()) +
  coord_flip() +
  labs(title = "24: ICH ~ Follow @ mri")

# for follow at discharge
avg1<- dat3.m %>% 
  group_by(follow2, variable) %>%
  filter( !(variable  %in%  c("Cerebellar tonsillar herniation","MLS [mm]", "Old stroke", "Old ICH",  "Uncal herniation (to which side) ","Transtentorial herniation ",
                              "Uncal herniation (to which side)","Transtentorial herniation","sf_hern", "Hypo_edema_C","Hypo_ICH_C") )) %>%
  #summarize(mean=mean(value,na.rm=TRUE), sd=sd(value,na.rm=TRUE)) no need sd (%)
  summarize(mean=mean(value,na.rm=TRUE))

table(loc_analyse.raw[loc_analyse.raw$sus_tent=='Both',]$follow2)

avg1$follow2<-as.integer(avg1$follow2);avg1$follow2<-as.factor(avg1$follow2)

ggplot(data=avg1[grep(pattern = "edema", avg1$variable),  ], aes(x=variable, y=mean, fill=follow2) )+
  geom_bar(stat="identity", position=position_dodge()) +
  coord_flip()+
  labs(title = "25: Edema ~ Follow @ discharge")

ggplot(data=avg1[grep(pattern = "ICH", avg1$variable),  ], aes(x=variable, y=mean, fill=follow2) )+
  geom_bar(stat="identity", position=position_dodge()) +
  coord_flip()+
  labs(title = "26: ICH ~ Follow @ discharge")
```